---
title: "Group-level analysis tutorial"
editor: source
format:
  html:
    toc: true
    standalone: true
    code-fold: true
---


```{r}
knitr::opts_chunk$set(echo = TRUE, eval=TRUE)
```

## Overview

We will analyze diagnostic and age associations in individuals with autism spectrum disorder and typically developing controls.
The data we will analyze are called fractional amplitude of low frequency fluctuations (fALFF).
The are available already processed from the Autism Brain Imaging Data Exchange (ABIDE) dataset in MNI space at 3mm resolution.

### Analysis goals and data input

```{r}
# install the latest version of the NIsim package, which includes a function to download the preprocessed ABIDE data.
# devtools::install_github('statimagcoll/NIsim', ref='master')

### LIBRARIES ###
library(RNifti)
library(parallel)
library(splines)
library(progress)
library(table1)
#library(pbj)
library(PCP)
```



### Outline of steps

1. Download ABIDE data
2. Data quality assurance
    * Create study level brain mask
    * Motion exclusions
    * Covariate exclusions
3. Group-level analysis
    * FEAT and `easythresh`
    * `Randomise`
    * `pbj` package in `R`
4. Reporting results

## Download ABIDE data

The `PCP` `R` package has access to the preprocessed ABIDE data. Here, the code downloads fALFF images, which is a measure derived from the rs-fMRI data.

```{r, eval=FALSE}
### DATA LOCATION ###
# SMN data folder
datadir = file.path(dirname(dirname(getwd())), 'data/abide')
derivative='falff'
# will be created by downloadABIDE
dbimagedir = file.path(datadir, 'neuroimaging/cpac/filt_global', derivative)
templatefile = file.path(datadir, 'neuroimaging/MNI152_T1_3mm.nii.gz')
# created by simSetup... Not used in this analysis
dbresimagedir = file.path(datadir, 'neuroimaging/cpac/alff_res')
# created below
maskfile = file.path(datadir, 'neuroimaging/cpac/mask.nii.gz')

# load in data and get directories
dat = ABIDE(datadir, derivatives=derivative)
# some data curation is done on download. Subset all pass on visual inspection and making sex and dx factors, remove rows with no file name
dat$imgname = paste(dat$file_id, paste0(derivative, '.nii.gz'), sep='_')
dat$files = file.path(dbimagedir, dat$imgname)
```


## Data quality assurance

```{r table1}
dat$sex = factor(dat$sex, levels=1:2, labels = c('Male', 'Female'))
label(dat$sex) = "Sex"
dat$dx_group = factor(toupper(dat$dx_group))
dat$site_id = factor(dat$site_id)
label(dat$age_at_scan) = "Age"
label(dat$func_mean_fd) = "Mean displacement"
label(dat$dx_group) = "Diagnosis"
table1(~ sex + func_mean_fd + age_at_scan | dx_group, data=dat, caption = "ABIDE analysis summary table.")
```


### Motion exclusions

```{r motion, fig.cap="Mean framewise displacement (FD) for each participant. The dashed line indicates the 0.5mm threshold we use to exclude participants with high motion."}
# plot motion
plot(dat$func_mean_fd, main="Mean relative frame displacement", bty='l')
# greater than half a millimeter
abline(h=0.5, lty=2, lwd=2)
# add number of exclusions as text in the top right corner of plot

# subset to participants less than half a milimeter mean framewise displacement
dat = dat[which(dat$func_mean_fd<0.5), ]
```


### Covariate exclusions

This is just a check to make sure we don't have any missing non-imaging variables.

```{r}

```



### Create study level brain mask

After downloading the data, we create a study specific mask based on the subset of the data where all subjects have coverage. We make a small number of exclusions to increase the coverage above 30,000 voxels at 3mm isotropic resolution. This first chunk of code creates a study-level mask by the intersection of all participants where their data are non-zero.
It the iteratively excludes the participant that will increase the study-level mask, until we reach a coverage over 30,000 voxels.

```{r}
imgs = simplify2array(readNifti(dat$files) )
# choose
# number of people with zeros at that location
inds=numeric()
ids = c()
subids = dat$sub_id
# number of voxels with no zeros
nnzero = 0
# iteratively removes subjects who will increase the mask the largest
while(nnzero<30000){
  voxSums = rowSums(imgs==0, dims=3)
  tab = as.data.frame(table(voxSums))
  nnzero=tab[1,2]
  # number of unique voxels for each subject
  uniquevox = apply(imgs, 4, function(img) sum(img==0 & voxSums==1) )
 
  inds = which.max(uniquevox)
  cat('\nIteration:\nsubject removed: ', paste(subids[inds], collapse=', '), '\nmask size is now ', nnzero+sum(uniquevox[inds]), '\nNumber of voxels added:', sum(uniquevox[inds]) )
  imgs = imgs[,,,-inds]
  subids = subids[-inds]
  nnzero=nnzero+sum(uniquevox[inds])
}
```


There are issues with the orientation of the images for a small number of participants. We can visualize them here to see what the problem is.

## Group-level analysis

### FEAT and `easythresh`

### `Randomise`

### `pbj` package in `R`

```{r}
# need age_at_scan in both models for testing nonlinear functions
form = paste0(" ~ sex + age_at_scan + func_mean_fd + func_fber + func_outlier + dx_group + site_id" )
formred = paste0(" ~ sex + age_at_scan + func_mean_fd + func_fber + func_outlier + site_id")
formredAge = paste0(" ~ sex + dx_group + func_mean_fd + func_fber + func_outlier + site_id")
#  weights for each subject. Can be a character vector
weights = "func_mean_fd"
mask = maskfile
output = paste0(tempfile(), '.rdata')
ncores = 24
```


## Reporting results

