---
title: "Group-level analysis tutorial"
editor: source
format:
  html:
    toc: true
    standalone: true
    code-fold: true
---


```{r}
knitr::opts_chunk$set(echo = TRUE, eval=TRUE)
```

# Overview

We will analyze diagnostic and age associations in individuals with autism spectrum disorder and typically developing controls.
The data we will analyze are called fractional amplitude of low frequency fluctuations [fALFF](http://www.sciencedirect.com/science/article/pii/S0165027008002458). They are computed from rs-fMRI data using the following steps:

1. The data are preprocessed as usual for rs-fMRI data.
2. fALFF is the power in the spectrum 0-0.08Hz divided by the power in the spectrum 0-0.25Hz.
3. fALFF is computed at every voxel in the image.

The fALFF data are available already processed from the Autism Brain Imaging Data Exchange (ABIDE) dataset in MNI space at 3mm resolution.

To analyze the data, we will look at associations of fALFF with diagnosis and age, controlling for other covariates.

$$
\text{fALFF}_i \sim \text{age}_i + \text{sex}_i + \text{diagnosis}_i + \text{motion covariates}_i + \text{site}_i
$$

```{r, message=FALSE}
# install the latest version of the NIsim package, which includes a function to download the preprocessed ABIDE data.
# devtools::install_github('statimagcoll/PCP', ref='master')
# devtools::install_github('statimagcoll/pbj', ref='master')

### LIBRARIES ###
library(RNifti)
library(parallel)
library(splines)
library(progress)
library(table1)
library(pbj)
library(PCP)
```



## Outline of steps

1. Download ABIDE data
2. Data quality assurance
    * Create study level brain mask
    * Motion exclusions
    * Covariate exclusions
3. Group-level analysis
    * FEAT and `easythresh`
    * `Randomise`
    * `pbj` package in `R`
4. Reporting results

# Download ABIDE data

The `PCP` `R` package has access to the preprocessed ABIDE data. Here, the code downloads fALFF images, which is a measure derived from the rs-fMRI data.

```{r, eval=TRUE}
## DATA LOCATION ##
# SMN data folder
datadir = file.path(dirname(dirname(getwd())), 'data/abide')
derivative='falff'
# will be created by downloadABIDE
dbimagedir = file.path(datadir, 'neuroimaging/cpac/filt_global', derivative)
templatefile = file.path(datadir, 'neuroimaging/MNI152_T1_3mm.nii.gz')
# created by simSetup... Not used in this analysis
dbresimagedir = file.path(datadir, 'neuroimaging/cpac/alff_res')
# created below
maskfile = file.path(datadir, 'neuroimaging/cpac/mask.nii.gz')
overlapfile = file.path(datadir, 'neuroimaging/cpac/overlap.nii.gz')

# load in data and get directories
dat = ABIDE(datadir, derivatives=derivative)
# some data curation is done on download. Subset all pass on visual inspection and making sex and dx factors, remove rows with no file name
dat$imgname = paste(dat$file_id, paste0(derivative, '.nii.gz'), sep='_')
dat$files = file.path(dbimagedir, dat$imgname)
```


# Data quality assurance

```{r table1}
dat$sex = factor(dat$sex, levels=1:2, labels = c('Male', 'Female'))
label(dat$sex) = "Sex"
dat$dx_group = factor(toupper(dat$dx_group))
dat$site_id = factor(dat$site_id)
label(dat$age_at_scan) = "Age"
label(dat$func_mean_fd) = "Mean displacement"
label(dat$dx_group) = "Diagnosis"
table1(~ sex + func_mean_fd + age_at_scan | dx_group, data=dat, caption = "ABIDE analysis summary table.")
```


## Motion exclusions

```{r motion, fig.cap="Mean framewise displacement (FD) for each participant. The dashed line indicates the 0.5mm threshold we use to exclude participants with high motion."}
# plot motion
plot(dat$func_mean_fd, main="Mean relative frame displacement", bty='l')
# greater than half a millimeter
abline(h=0.5, lty=2, lwd=2)
# add number of exclusions as text in the top right corner of plot

# subset to participants less than half a milimeter mean framewise displacement
dat = dat[which(dat$func_mean_fd<0.25), ]
```


## Covariate exclusions

This is just a check to make sure we don't have any missing non-imaging variables. One additional participant is excluded.

```{r}
dat = dat[rowMeans(!is.na(dat[,c("sex", "age_at_scan", "func_mean_fd", "func_fber", "func_outlier", "dx_group", "site_id")]))==1,]
```



## Create study level brain mask { .tabset .tabset-fade}

After downloading the data, we create a study specific mask based on the subset of the data where all subjects have coverage. We make a small number of exclusions to increase the coverage above 30,000 voxels at 3mm isotropic resolution. This first chunk of code creates a study-level mask by the intersection of all participants where their data are non-zero.
It the iteratively excludes the participant that will increase the study-level mask, until we reach a coverage over 30,000 voxels.

```{r}
imgs = simplify2array(readNifti(dat$files) )
overlap = rowMeans(imgs!=0, dims=3)
overlapImg = readNifti(dat$files[1])
overlapImg[,,] = overlap[,,]
writeNifti(overlapImg, file=overlapfile)

# choose
# number of people with zeros at that location
inds=numeric()
ids = c()
subids = dat$sub_id
# number of voxels with no zeros
nnzero = 0
# iteratively removes subjects who will increase the mask the largest
while(nnzero<30000){
  voxSums = rowSums(imgs==0, dims=3)
  tab = as.data.frame(table(voxSums))
  nnzero=tab[1,2]
  # number of unique voxels for each subject
  uniquevox = apply(imgs, 4, function(img) sum(img==0 & voxSums==1) )
 
  inds = which.max(uniquevox)
  cat('\nIteration:\nsubject removed: ', paste(subids[inds], collapse=', '), '\nmask size is now ', nnzero+sum(uniquevox[inds]), '\nNumber of voxels added:', sum(uniquevox[inds]) )
  imgs = imgs[,,,-inds]
  subids = subids[-inds]
  nnzero=nnzero+sum(uniquevox[inds])
}
```

Create the mask from the overlap.

```{r}
badDat = dat[! dat$sub_id %in% subids,]
dat = dat[ dat$sub_id %in% subids,]

# now create the mask
mask = readNifti(dat$files[1])
imgs = simplify2array(readNifti(dat$files) )
# mask actually still has pretty low coverage
mask[,,] = 0
mask[,,] = apply(imgs>0, 1:3, all)
writeNifti(mask, maskfile)
nvox = sum(mask)
ulimit = quantile(apply(imgs, 4, function(x) x[mask!=0]), 0.9)
rm(imgs)
```

## Subjects with good coverage


Visualize a slice of the data for all participants included in the analysis.

::: {.panel-tabset}

```{r, fig.height=8, fig.width=12, results='asis'}
# hardcoded at 90 percentile
ulimit = 126.95
niftis = readNifti(dat$files)
#image(niftis[[5]], index=45, crop=FALSE)
#image(niftis[[5]], index=45, crop=TRUE, title='test', fg='white', bg='black')
# loop through and use tabset for each figure
# create 10 image within each figure
it = 0
par(mfrow=c(2,5), mar=c(0,0,2,0))
for(i in 1:length(niftis)){
  if(i%%10==1){ it = it+1; cat('\n\n### P', it, ' \n')}
  image(niftis[[i]], index=45, lo=FALSE, limits=c(0,ulimit), crop=FALSE)
  mtext(dat$sub_id[i], outer=FALSE)
}
```

:::



## Subjects with bad coverage

There are issues with the orientation of the images for a small number of participants. We can visualize them here to see what the problem is.

```{r, fig.cap="Axial slice (z=45) of fALFF data for all participants excluded from the analysis due to poor coverage.", fig.height=8, fig.width=12, results='asis'}
niftis = readNifti(badDat$files)
# visualize the ones that had bad coverage
par(mfrow=c(2,5), mar=c(0,0,2,0))
for(i in 1:length(niftis)){
  image(niftis[[i]], index=45, lo=FALSE, limits=c(0,ulimit), crop=FALSE)
  mtext(badDat$sub_id[i], outer=FALSE)
}
```


# Group-level analysis

```{r}
analysisDir = file.path(getwd(), "abide_results")
# need age_at_scan in both models for testing nonlinear functions
form = paste0(" ~ sex + age_at_scan + func_mean_fd + func_fber + func_outlier + dx_group + site_id" )
formred = paste0(" ~ sex + age_at_scan + func_mean_fd + func_fber + func_outlier + site_id")
formredAge = paste0(" ~ sex + dx_group + func_mean_fd + func_fber + func_outlier + site_id")
mask = maskfile
# output Rdata file
output = paste0(analysisDir, 'abideModels.rdata')
ncores = 4
```


## `fsl_glm`: OLS regression

We can use FSL's FEAT to do a group-level analysis using a simple GLM (least squares approach). Here we create a text file with the list of images, and text files with the design matrix and contrasts for diagnosis and age.


```{r}
# feat setup files directory
featDir = file.path(analysisDir, 'feat')
dir.create(featDir, recursive=TRUE, showWarnings = FALSE)

# write text file with all iamges
write.table(dat$files, file = file.path(featDir, 'images.txt'), row.names=FALSE, col.names=FALSE, quote=FALSE)

# create design file for feat
# save contrast names
# create design matrix without the intercept for feat and save it as a text file
design = model.matrix(as.formula(form), data=dat)
write.table(design, file=file.path(featDir, 'design.txt'), row.names=FALSE, col.names=FALSE, quote=FALSE)

# create contrasts for dx and age
contrasts = matrix(0, nrow=2, ncol=ncol(design))
colnames(contrasts) = colnames(design)
rownames(contrasts) = c('dx', 'age')
contrasts[1, which(colnames(contrasts)=='dx_groupHC')] = 1
contrasts[2, which(colnames(contrasts)=='age_at_scan')] = 1
# save contrasts
write.table(contrasts, file=file.path(featDir, 'contrasts.txt'), row.names=FALSE, col.names=FALSE, quote=FALSE)
getwd()
```


```{bash, eval=FALSE}
# output directory
featDir=$(pwd)/abide_results/feat
# merge images to 4d file
fslmerge -t $featDir/falff_outcome.nii.gz $(cat $featDir/images.txt)
# convert text file to .mat format for fsl
Text2Vest $featDir/design.txt $featDir/design.mat
# convert contrast text file to fsl format
Text2Vest $featDir/contrasts.txt $featDir/design.con
# mask file
mask=$(pwd)/../../data/abide/neuroimaging/cpac/mask.nii.gz
# run glm command
fsl_glm -i $featDir/falff_outcome.nii.gz -d $featDir/design.mat -c $featDir/design.con -m $mask \
-o $featDir/pe.nii.gz --out_z=$featDir/zstat.nii.gz  --out_t=$featDir/tstat.nii.gz --out_p=$featDir/pstat.nii.gz --out_cope=$featDir/cope.nii.gz
```


## `flameo`: random effects model with subject-specific error variance

FSL's FLAME model accounts for differences in variance among participants. It also can account for repeated measurements among participants. Might be helpful to review slides 12-16 in the [group-level analysis notes](4group-level_analysis/pptx/group-level_analysis.pdf).

## `randomise`: permutation approach

## `pbj` package in `R`: robust variance estimation

First, try to replicate the GLM results from fsl

```{r}
# DX analysis
abideDXglm = lmPBJ(dat$files, form=form, formred=formred, mask=mask, data=dat, template = templatefile, robust=FALSE, HC3 = FALSE, transform='t')
abideDXglm
# effect size for diagnosis effect
write.statMap(abideDXglm, outdir=file.path(analysisDir, 'dxGLM'), statMethod = 'Z')
```


This analysis uses the default settings, which includes a robust covariance estimator.

```{r}
# DX analysis
abideDX = lmPBJ(dat$files, form=form, formred=formred, mask=mask, data=dat, template = templatefile, transform='t')
abideDX
```



```{r, fig.height=30, fig.width=50}
image(abideDX, cft_p=0.05, crop=TRUE, index = 18:47)
write.statMap(abideDX, outdir=file.path(analysisDir, 'dx'), statMethod = 'Z')
```

Load interactively with `fsleyes`. Open also with MNI 1mm template for better resolution viewing. `fsleyes` can overlay them correctly.

```{bash, eval=FALSE}
fsleyes /Users/vandeks/Library/CloudStorage/Box-Box/SMN/data/abide/neuroimaging/MNI152_T1_3mm.nii.gz /Users/vandeks/Library/CloudStorage/Box-Box/SMN/4group-level_analysis/group-level_tutorial/abide_results/dx/stat_p.nii.gz
```


The same approach but testing the age effect.

```{r}
# Age analysis
abideAge = lmPBJ(dat$files, form=form, formred=formredAge, mask=mask, data=dat, template = templatefile, transform='t')
```

```{r, fig.height=30, fig.width=50}
image(abideAge, cft_p=0.05, crop=TRUE, index = 18:47)
write.statMap(abideAge, outdir=file.path(analysisDir, 'age'))
write.statMap(abideAge, outdir=file.path(analysisDir, 'age'), statMethod = 'S')
```

# Reporting results

