---
title: "Despicable Me group-level analysis"
editor: source
format:
  html:
    toc: true
    standalone: true
    code-fold: true
---


```{r}
knitr::opts_chunk$set(echo = TRUE, eval=TRUE)
```


# Introduction

We will perform a group-level analysis of the *Despicable Me* task.

# Setup

```{r}
sites = read.csv('/Users/vandeks/Library/CloudStorage/Box-Box/SMN/3first-level_analysis/homework/first-level_analysis/subject_assignment.csv')
sites = sites[ ! sites$student %in% c('Yisu', 'Shuai', 'Laura') & !is.na(sites$student),]
# list available feat directories
feats = sapply(strsplit(system("ls -d /Users/vandeks/Library/CloudStorage/Box-Box/SMN/data/RBC/HBN_BIDS/*/*/func/*bold.feat", intern = TRUE), "/"), function(x) x[[11]])
suppressWarnings(feats <- sapply(sites$subject, function(x)
  system(
    paste0("ls -d /Users/vandeks/Library/CloudStorage/Box-Box/SMN/data/RBC/HBN_BIDS/", x, "/*/func/*bold.feat 2>/dev/null"),
    intern = TRUE) ))
feats = unlist(feats)
sites$feat = NA
sites$feat[ match(names(feats), sites$subject)] = feats
```


## Merge in demographic variables

For the group-level analyses, we are going to look at how some demographic features are associated with BOLD response during the task.

* Merge in the demographic variables. They are located here `data/RBC/HBN_BIDS/study-HBN_desc-participants.tsv` in tab delimited format.

## Motion QC

The motion files are included in the feat directories in the file `mc/prefiltered_func_data_mcf_rel_mean.rms`.

* Add them to the sites data frame and create a plot to decide on a threshold for exclusion.
* Create a list of participants to exclude based on the visual registration QA.
* Make the exclusions and create a Table 1 with the motion information.

```{r}
# get mean relative displacement from feat directories and add to sites dataframe

# create a plot with the motion QA and decide on a threshold for exclusion

# create list of participants to exclude based on our visual registration QA
regExclude = c()
```


## Study-level mask

* Create an "overlap" image that shows the proportion of participants with data at each voxel based on the mask in their FEAT folder. Visualize the overlap image here with the MNI in the background.
* Create a study-level mask that includes voxels with 100% of data (i.e. every participant has data).
* Write a one-sentence comment on how the coverage looks.



## Group level model


### FSL OLS and FLAMEO

Fit an OLS model and FLAME model with the following design:
$$
Y_i(v) = \beta_0(v) + \beta_1(v)*\text{age}_i + \beta_2(v)*\text{sex}_i + \beta_3(v)*\text{motion}_i,
$$

where $Y_i$ the the "cope" file for the difference between positive and negative emotional states.
For age, use the centered age (subtract the mean age from each participant's age).


For the group level contrasts, create three:

1. The mean effect across all participants.
2. The association with age (positive correlation with age).
3. The association with sex.

* For each contrast, describe the interpretation of the effect.

**Hint:** For the OLS model, you can use the group-level tutorial that I presented in class. For the FLAME model, you can use the `flameo` function. You will need to pass the `varcope` images in addition to the `cope` images. The `varcope` images are also located in the feat directories.


### `pbj` package

* Use the `pbj` package to fit the same model for the age and sex effects.



### Visualizations

* Threshold the statistical images for the age effect showing positive and negative regions at $Z\ge 1.96$ and create an image here with a particular slice for the three models (OLS, FLAMEO, pbj) that shows some of the difference between the age effects side-by-side. **Note:** I haven't done this analysis yet. If it looks like there isn't much showing, up, then let me know and we can discuss further.



## Spatial extent inference


### easythresh

* Use the `easythresh` function from FSL to perform spatial extent inference for the age effect on the FLAME output. Use a cluster-forming threshold of $p<0.001$.
* Create a cluster table for the results.

### pbj

This will take a little while longer to run because it is using bootstraps to compute $p$-values.

* Use the `pbj` package to perform spatial extent inference for the age effect. Use a cluster-forming threshold of $p<0.001$ and a family-wise error rate of $\alpha=0.05$. How many clusters are significant?
* Create a cluster table for the results.

## Reporting age effects

For this section, report the results from the FLAME or pbj analysis (you can choose which one) for the age effect.

### Methods

Describe the methods you used to analyze the results 1-2 paragraphs including a description of the first-level analysis and group-level analysis. It should be described well enough that someone could repeat your analyses.

### Results

Describe the results of your analyses on one paragraph, including a final figure showing the significant clusters from the spatial extent inference for the age effect. Include a table with the cluster information. You should interpret the location of the result and why you think it is relevant to the task and the contrast.

